RCloud: Integrating Exploratory Visualization, Analysis and Deployment

### 2014-03-30 paper.tex
notebook figure and section

%% Uncomment below to include a teaser figure.
\teaser{
  \centering
  \includegraphics[width=.3\linewidth]{fig/screenshots/wdcplotsmall.jpg}
  %% \includegraphics[width=.9\linewidth]{figs/teaser.jpg}
  % \vspace{-.5em}
  \caption{Left: RCloud integration of HTML5 visualization libraries such
    as D3~\cite{Bostock:2011:DDD} and Crossfilter~\cite{Square:2014:CFM}.
  }
% \vspace{-.5em}
}

>>>>

%% Uncomment below to include a teaser figure.
\teaser{
  \centering
  \includegraphics[width=.3\linewidth]{fig/screenshots/wdcplotsmall.jpg}
  %% \includegraphics[width=.9\linewidth]{figs/teaser.jpg}
  % \vspace{-.5em}
  \caption{RCloud provides infrastructure for data science teams so
    that R \emph{hackers} and \emph{scripters}~\cite{Kandel:2012:EDA} can target modern
    interactive visualization technologies such as
    D3~\cite{Bostock:2011:DDD}, dc.js~\cite{DCJS:2014:DC} and
    Crossfilter~\cite{Square:2014:CFM} (left), while at the same time
    enabling seamless deployment and dashboarding for
    \emph{application users} (right)
  }
% \vspace{-.5em}
}

### 2014-03-30 system.tex
update
notebook figure and section

\section{The System}

@@

Notebooks as Github gists. 
%
We needed version control over a directory, gist provides that,
wrapped over an HTTP interface.
%
Mainly a tactical decision to save time.
%
Nevertheless, availability of notebook data as (barely) plain text had good side
effects, like ease to build search. 
%
Expect same with suggest.

@@

>>>>

\section{The System\label{sec:system}}

@@

\begin{figure*}
\includegraphics[width=\linewidth]{fig/notebook/notebook.pdf}
\caption{\label{fig:notebook}An RCloud notebook is a sequence of
  \emph{cells}, each a snippet of source in R or Markdown. Notebooks
  are stored in GitHub \emph{gists}, which provide the version-control
  primitives necessary for RCloud.}
\end{figure*}

The basic unit of operation in RCloud is a \emph{notebook}. RCloud
notebooks hold a sequence of \emph{cells}, each of which represents a
snippet of code or hypertext in Markdown. This is not a novel idea;
executable documents structured in this way are present in a wide
variety of present software systems, from Mathematica to IPython to
Sage and others.

One of the main differences between RCloud and other systems is the
notion that notebooks are ``always deployed''. This has a variety of
advantages which we discuss \todo{where?}. On the other hand, there
are a variety of situations that might ask for a specific version of a
notebook. Rather than have the notebook designer decide ahead of time
which versions should be preserved, we choose to expose
\emph{transparent} versioning. This is similar to models like
Jankun-Kelly et al.'s p-set calculus \todo{cite} and VisTrails's
version tree \todo{cite}, where every change in the state of the
system is tracked.

Implementing the entire version-control subsystem would be a large
undertaking. Instead, we build the infrastructure on top of Github's
\emph{gists}~\todo{cite}. Github offers a HTTP interface
for creating simplified git repositories (the main limitation is the
use of text-only files and a single directory). The GitHub web-service
API provides much of the infrastructure we need for the versioning
portion of the storage back end: access to previous versions,
comments, starring, and forking.

Using GitHub was mainly a tactical decision to reduce our
time-to-prototype. At the same time, it had the fortuitous decision of
forcing our backend to communicate with separate processes over HTTP
and JSON. This had the serendipitous advantage of quickly enabling
full search over RCloud notebooks (built using Apache SOLR).
%
This perspective has implications for
systems-based research and prototyping in visual analytics.
%
In order for small teams to efficiently explore ideas in the present
software ecosystem, it pays off to use as many existing solutions as
possible.
%
Even though it involved learning a variety of standards that are not
directly related to visual analytics, our early decision to use the
GitHub gist API has been quite fruitful. We expect similar payoffs to
happen for related projects as well.

%% %
%% We needed version control over a directory, gist provides that,
%% wrapped over an HTTP interface.
%% %
%% Mainly a tactical decision to save time.
%% %
%% Nevertheless, availability of notebook data as (barely) plain text had good side
%% effects, like ease to build search.
%% %
%% Expect same with suggest.

@@

\subsection{Executing R in a web browser}

As we reiterated in 
One of the design goals of the user experience in RCloud
